|

### 2021-08-15 dissertation/msc_dissertation.Rmd
Cleaning up the code

## Outcomes of running the 1,000 models
dat_models = data.frame(fread(here::here("data", "4_model_outcome_data.csv")))

# 
dat_models_specifications = data.frame(fread(here::here("data", "4_model_outcome_data_specifications.csv")))

@@

# Adjusting class for proper visualisation
dat_models$below_alpha = factor(dat_models$below_alpha)

@@

# Order by odds ratio, low to high
analysed_specifications = dat_models_specifications[order(dat_models_specifications$estimate_oddsratio), ]

# Assign identifier
analysed_specifications$id = factor(seq(1:nrow(analysed_specifications)))

# Preparing specification matrix i.e., the bottom part of the SCA plot
## Subset covariates and statistics
prepare_long_df = analysed_specifications[ , c(4:18, 20, 21, 24)]

## FALSE ro NA to better subset
prepare_long_df[prepare_long_df == FALSE] = NA

## Wide to long format
long_df = melt(setDT(prepare_long_df), id.vars = c("id", "below_alpha"), variable.name = c("estimate_oddsratio"))

## Exclude NA cases
long_df = long_df[complete.cases(long_df), ]

## Exclude outcome variable
long_df = long_df[!grepl("estimate_oddsratio", long_df$estimate_oddsratio), ]
colnames(long_df) = c("id", "below_alpha", "covariate", "x")

## Exclude value column
long_df = long_df[ , c("id", "below_alpha", "covariate")]

## order covariate effects low to high, based on the 4_covariate_effect_data.csv dataset
impact_df = dat_covariates[order(dat_covariates$estimate_oddsratio), ]

long_df$covariate = factor(long_df$covariate,
                           levels = c("club", "victories", "weight_kg", "player_cards_received", "ties", "league_country", "specific_pos", "age_yrs", "ref", "ref_cards_assigned", "goals", "height_cm", "games", "ref_country", "player"),
                           labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "Lg. country", "Position", "Age", "Ref", "Cards assig.", "Goals", "Height", "Games", "Ref country", "Player"))

# build plot
top = analysed_specifications %>% 
  ggplot(aes(as.numeric(id), estimate_oddsratio)) +
  geom_point(aes(colour = below_alpha), size = 0.3, alpha = 0.5) +
  geom_ribbon(aes(ymin = ci_lower_oddsratio, ymax = ci_upper_oddsratio), alpha = 0.3) +
  geom_hline(yintercept = 1, color = "black") +
  scale_color_manual(values = c("red", "black")) +
  scale_x_discrete(name = "", expand = c(0.01, 0)) +
  scale_y_continuous(name = "Odds ratio", breaks = c(seq(0.9, 1.7, 0.1)), limits = c(0.89, 1.71), expand = c(0, 0)) +
  labs(title = "Specification curve", subtitle = "N = 1,000; 95% CI") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        plot.margin = unit(c(0.5, 1, 0, 1), "cm"),
        legend.position = "top",
        legend.title.align = 1,
        legend.text = element_text(size = 11),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
        legend.box.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic", margin = margin(0, 0, 20, 0)),
        text = element_text(size = 10)) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 1), reverse = TRUE))

bottom = long_df %>% 
  filter(id != "1") %>%
  ggplot(aes(x = id, y = covariate)) +
  geom_tile(aes(fill = below_alpha), width = 0.5, height = 0.5, color = "white") +
  scale_fill_manual(values = c("red", "black")) +
  scale_x_discrete(name = "Specifications", expand = c(0.01, 0)) +
  scale_y_discrete(name = "Covariates", expand = c(0.03, 0)) +
  theme_classic() +
  theme(legend.position = "none", 
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(0, 1, 0.5, 1), "cm"),
        text = element_text(size = 10))

@@

## Results space
     Shown in Figure \ref{fig:rain_cloud_plot}. The horizontal axis shows the odds rations. From top to bottom the three component show the probability density distribution, the box plot and the raw data, respectively. The dashed line indicates the original results of *team 23*, which has also been the median outcome of all analytical strategies in @silberzahn2018. Based on the raw data points, the results space can be defined as the interval between 1.081 and 1.383. The 1rd quartile   \newline

@@

vibration_of_effect = dat_models %>%
  filter(estimate_oddsratio > 0 & estimate_oddsratio <= 4) %>% 
  ggplot(aes(x = "", y = estimate_oddsratio)) +
  geom_flat_violin(aes(fill = ""), trim = FALSE, colour = "dark grey", fill = "dark grey") +
  geom_point(aes(x = 0.6, y = estimate_oddsratio, colour = below_alpha),
             position = position_jitter(width = .03, seed = 123), size = 3, shape = 20, alpha = 0.3, stroke = 0) +
  geom_boxplot(aes(x = 0.81, y = estimate_oddsratio), alpha = 0.5, width = 0.1, colour = "black") +
  geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
  annotate("text", x = 1.516, y = 1.375, label = "Silberzahn et al. (2018) \nMedian OR", color = "black") +
  scale_color_manual(values = c("red", "black")) +
  scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
  labs(title = "Results space defined through covariate specification", subtitle = "N = 1,000; Odds ratio (OR)") +
  coord_flip() +
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
        legend.position = "top",
        legend.justification = "left",
        legend.title.align = 1,
        legend.text = element_text(size = 11),
        legend.direction = "vertical",
        legend.title = element_blank(),
        legend.margin = margin(b = -0.8, unit = "cm"),
        legend.box.margin = margin(b = -0.8, unit = "cm"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        text = element_text(size = 10)) +
  guides(colour = guide_legend(override.aes = list(alpha = 1), reverse = TRUE))

@@

## Specified covariate effects
     \newline

@@

covariate_effects = dat_covariates %>%
  ggplot(aes(x = reorder(x = impact_names, X = impact_coef), impact_coef)) +
  geom_point(aes(color = below_alpha)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper, color = below_alpha), width = 0.1) +
  scale_color_manual(values = c("red", "black")) +
  scale_y_continuous(name = "Estimates", labels = scales::comma_format(accuracy = 0.05), breaks = c(seq(-0.15, 0.15, 0.05)), limits = c(-0.15, 0.15)) +
  labs(title = "Specified covariate effects", subtitle = "N = 1,000; 95% CI", x = "Covariates") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
        legend.position = "top",
        legend.justification = "left",
        legend.title.align = 1,
        legend.text = element_text(size = 11),
        legend.direction = "vertical",
        legend.title = element_blank(),
        legend.margin = margin(l = 0.1, b = -0.729, unit = "cm"),
        legend.box.margin = margin(l = 0.1, b = -0.729, unit = "cm"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        text = element_text(size = 10)) +
  guides(colour = guide_legend(reverse = TRUE))

>>>>

## Outcomes of running the 1,000 models
dat_models_specifications = data.frame(fread(here::here("data", "4_model_outcomes_specifications_merged.csv")))

@@

# Adjusting class for proper visualisation
dat_models_specifications$below_alpha = factor(dat_models_specifications$below_alpha)

@@

# Order by odds ratio, low to high and assign identifier
analysed_specifications = dat_models_specifications[order(dat_models_specifications$estimate_oddsratio), ]
analysed_specifications$id = factor(seq(1:nrow(analysed_specifications)))

# Preparing the bottom part of the SCA plot by transforming wide to long
## Subset covariates and relevant statistics
prepare_long_df = analysed_specifications %>% select(specific_pos:games, id)

## transform wide to long
long_df = melt(data.table(prepare_long_df), "id", variable.names = "covariate")
long_df = long_df[long_df$value == TRUE, c("id", "variable")]

## statistics are still missing (odds ratio and significance)
id_oddsratio_below_alpha = analysed_specifications %>% select(id, estimate_oddsratio, below_alpha)

## Adding (or joining) the statistics to the long data frame
long_df = data.frame(left_join(long_df, id_oddsratio_below_alpha, by = "id"))

# Assign proper names for visualisation
long_df$variable = factor(long_df$variable,
                           levels = c("club", "victories", "weight_kg", "player_cards_received", "ties", "league_country", "specific_pos", "age_yrs", "ref", "ref_cards_assigned", "goals", "height_cm", "games", "ref_country", "player"),
                           labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "Country", "Position", "Age", "Referee", "Cards assig.", "Goals", "Height", "Games", "Ref. country", "Player"))

# build plot
top = analysed_specifications %>% 
  filter(estimate_oddsratio < 4) %>% 
  ggplot(aes(as.numeric(id), estimate_oddsratio)) +
  geom_point(aes(colour = below_alpha), size = 0.3, alpha = 0.5) +
  geom_ribbon(aes(ymin = ci_lower_oddsratio, ymax = ci_upper_oddsratio), alpha = 0.3) +
  geom_hline(yintercept = 1, color = "black") +
  scale_color_manual(values = c("red", "black")) +
  scale_x_discrete(name = "", expand = c(0.01, 0)) +
  scale_y_continuous(name = "Odds ratio", breaks = c(seq(0.9, 1.7, 0.1)), limits = c(0.89, 1.71), expand = c(0, 0)) +
  labs(title = "Specification curve", subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = "")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        plot.margin = unit(c(0.5, 1, 0, 1), "cm"),
        legend.position = "top",
        legend.title.align = 1,
        legend.text = element_text(size = 11),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
        legend.box.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic", margin = margin(0, 0, 20, 0)),
        text = element_text(size = 10)) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 1), reverse = TRUE))

bottom = long_df %>%
  ggplot(aes(x = id, y = reorder(x = variable, X = as.numeric(estimate_oddsratio)))) +
  geom_tile(aes(fill = below_alpha), width = 0.5, height = 0.5, color = "white") +
  scale_fill_manual(values = c("red", "black")) +
  scale_x_discrete(name = "Specifications", expand = c(0.01, 0)) +
  scale_y_discrete(name = "Covariates", expand = c(0.03, 0)) +
  theme_classic() +
  theme(legend.position = "none", 
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(0, 1, 0.5, 1), "cm"),
        text = element_text(size = 10))

@@

## Results space
     Shown in Figure \ref{fig:rain_cloud_plot}. The horizontal axis shows the odds rations. From top to bottom the three component show the probability density distribution, the box plot and the raw data, respectively. The dashed line indicates the original results of *team 23*, which has also been the median outcome of all analytical strategies in @silberzahn2018. Based on the raw data points, the results space can be defined as the interval between 1.081 and 1.383. The 1st quatile equals to 1.206, the median equals to 1.248 and the 3rd quartile equals to 1.293. Therefore, as also evidenced by the probability density distribution, 50% of the lay withing a very narrow interval (between 1st and 3rd quartile). The outcome of *team 23* (1.31) is still included in the interval, but it is outside the middle 50% of the data.

 The probability density distribution also has an interesting shape, having two peaks. This suggests a particular covariate combination structure causing two focal points. Based on this graph alone though it is not possible to decipher the underlying structure though. What also seem to be more apparent in this graph than in the specification curve is that the non-significant values also seem to revolve round to focal points. One being around 1.13 and the other around 1.19. This also suggest some influential covariate or covariate structure that is cause non-significance.

 Finally, for transparency purpose, it also needs to be mentioned that eight points were excluded from the visualisation. Four of them were "infinite", one effectively approaching infinity, one being in the lowest million range, one being 150, and a final being 4.6. I specifically chose not to include the latter one in the graph as it would despite is relatively close numerical proximity skewed the distribution being at least a visual outlier. \newline

@@

vibration_of_effect = dat_models_specifications %>%
  filter(estimate_oddsratio < 4) %>% 
  ggplot(aes(x = "", y = estimate_oddsratio)) +
  geom_flat_violin(aes(fill = ""), trim = FALSE, colour = "dark grey", fill = "dark grey") +
  geom_point(aes(x = 0.6, y = estimate_oddsratio, colour = below_alpha),
             position = position_jitter(width = .03, seed = 123), size = 3, shape = 20, alpha = 0.3, stroke = 0) +
  geom_boxplot(aes(x = 0.81, y = estimate_oddsratio), alpha = 0.5, width = 0.1, colour = "black") +
  geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
  annotate("text", x = 1.516, y = 1.375, label = "Silberzahn et al. (2018) \nMedian OR", color = "black") +
  scale_color_manual(values = c("red", "black")) +
  scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
  labs(title = "Results space defined through covariate specification", subtitle = paste0("N = ", nrow(dat_models_specifications), ", Odds ratio (OR)", collapse = "")) +
  coord_flip() +
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
        legend.position = "top",
        legend.justification = "left",
        legend.title.align = 1,
        legend.text = element_text(size = 11),
        legend.direction = "vertical",
        legend.title = element_blank(),
        legend.margin = margin(b = -0.8, unit = "cm"),
        legend.box.margin = margin(b = -0.8, unit = "cm"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        text = element_text(size = 10)) +
  guides(colour = guide_legend(override.aes = list(alpha = 1), reverse = TRUE))

@@

## Specified covariate effects
     The final graph shows the specified covariate effects. On the horizontal axis are the covariates, on the vertical axis are the estimates. The covariates are sorted from negative to positive impact. Black indicates a significant effect, red non-significant. The error bars represent the 95% CI. While four covariates are significant, their effects seem relatively weak, ranging from -0.1 to 0.05, only. Interesting is also which covariates are significant. The *team 23* analysis included league country, position, referee and player as covariates.  \newline

@@

covariate_effects = dat_covariates %>%
  ggplot(aes(x = reorder(x = impact_names, X = impact_coef), impact_coef)) +
  geom_point(aes(color = below_alpha)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper, color = below_alpha), width = 0.1) +
  scale_color_manual(values = c("red", "black")) +
  scale_y_continuous(name = "Estimates", labels = scales::comma_format(accuracy = 0.05), breaks = c(seq(-0.15, 0.15, 0.05)), limits = c(-0.15, 0.15)) +
  labs(title = "Specified covariate effects", subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = ""), x = "Covariates") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
        legend.position = "top",
        legend.justification = "left",
        legend.title.align = 1,
        legend.text = element_text(size = 11),
        legend.direction = "vertical",
        legend.title = element_blank(),
        legend.margin = margin(l = 0.1, b = -0.729, unit = "cm"),
        legend.box.margin = margin(l = 0.1, b = -0.729, unit = "cm"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic"),
        text = element_text(size = 10)) +
  guides(colour = guide_legend(reverse = TRUE))
